### 2016-11-05 articles_of_association_english_french.tex
articles updated after general assembly 2016

%!TEX program = xelatex

\documentclass[12pt,a4paper,oneside]{article}
\usepackage{paracol}
%\usepackage{fontspec}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage[margin=2cm]{geometry}
\usepackage{navigator}
\usepackage[utf8]{inputenc}

\newcounter{art}
\addtocounter{art}{1}

\newcommand{\english}{    \switchcolumn[0]\noindent}
\newcommand{\french}{    \switchcolumn[1]\noindent}
\renewcommand{\thesubsubsection}{Art. \arabic{art} }
\renewcommand{\thesubsection}{\arabic{subsection}}

%\setmainfont{Gibson Light}
%\newfontfamily\semibold{Gibson}

\setcounter{section}{1}
\begin{document}

\begin{center}
	\includegraphics{octanis_org_logo_large}
\end{center}

\begin{paracol}{2}\sloppy

\english
	\section*{Articles of Association}

% =================================

\english
	\subsection{Name, Domicile and Purpose}
	\subsubsection{Name}\stepcounter{art}
	In the name of
	\begin{center}
		\textbf{Octanis}
	\end{center}
	a non-profit association of legal status according to articles 60 et  seq.  of  the  Swiss  Civil  Code is established.

\english
	\subsubsection{Domicile}\stepcounter{art}
	The associations headquarters are in Lausanne.

\english
	\subsubsection{Purpose}\stepcounter{art}
	The association has the following aims:

	\begin{enumerate}[(a)]

	\item Provide a place for curious and creative minds to embark on exploratory missions around the world.
	\item Distill our mission's findings into pragmatic solutions and inventions, accessible to everyone. 
	\item Promotion of the concept of rapid prototyping and design thinking by developing missions, prototypes, products, services with these and similar approaches.
	\item Embracing entrepreneurship by letting developed ideas be spun off into companies or be directly distributed through the association.
	\item Giving back to the open source community by releasing software and documentation where possible.
	\item Provide activities on the campuses of EPFL and other universities and schools.

	\end{enumerate}

% =================================

\english
	\subsection{Membership}

\english
	\subsubsection{Code of Conduct}\stepcounter{art}
	Every member has an obligation to know and understand the guidelines contained within the Code of Conduct.

\english
	\subsubsection{Membership Conditions}\stepcounter{art}
	Any person who is willing to participate and contribute to the associations cause can become a member.

\english
	\subsubsection{Membership Duration}\stepcounter{art}
	Memberships are automatically renewed every year.

\english
	\subsubsection{Ceasing of Membership}\stepcounter{art}
	A member can cancel his membership at any time by sending an email to the vice president. Exclusion of members can be decided by the Executive Committee in cases of violation of the Articles of Association and/or Code of Conduct and can be executed with immediate effect.

\english
	\subsubsection{Membership Benefits}\stepcounter{art}
	Members have access to the full workshop of the association and are allowed to use 3D printing, manual/reflow soldering, parts, materials and various other tools to create their prototypes.

\english
	\subsubsection{Membership Fees}\stepcounter{art}

	\begin{enumerate}[(a)]
	\item Member outside Committee: \\ 50.- / year
	\item Executive Committee Member: \\ 100.- / year

	\end{enumerate}

\english
	\subsubsection{Voting Rights}\stepcounter{art}
	 Members outside the committee and Executive Committee members have equal voting rights. Passive members have no voting rights.
	 	\vspace{0.5cm}

\english
	\subsubsection{Honorary Membership}\stepcounter{art}
	Members who have demonstrated outstanding and selfless committment to the cause of the association may be awarded an honorary membership by the General Assembly. The honorary membership does not offer additional voting rights.

\english
	\subsubsection{Passive Membership}\stepcounter{art}
	Members observing the associations activities without participating actively fall into the category of a Passive Member. Passive Members pay no membership fee but are also not permitted to access tools, materials or any of the other facilities the association provides. Passive members have no voting rights.
	\vspace{0.6cm}

% =================================

\english
	\subsection{Organisation}

\english
	\subsubsection{Bodies}\stepcounter{art}

	\begin{enumerate}[(a)]
	\item General Assembly
	\item Executive Committee
	\end{enumerate}

\english
	\subsubsection{General Assembly}\stepcounter{art}
	The General Assembly is the associations supreme authority. It is composed of all the members and is held annually. An Extraordinary General Assembly which discusses a specific topic or issue can be called for at any time.

	The Committee shall send notice to members at least a week in advance via email including the proposed agenda.

	Proposals by members must be submitted to the president at least a week in advance.

	The agenda of the annual General Assembly includes:
	\begin{enumerate}[(a)]
	\item Election of Committee members and auditors.
	\item Approval and changes to the Articles of Association.
	\item Approval of the reports and accounts, report of treasurer.
	\item Approval of the budget.
	\item Assessment of activities.
	\item Proposals.
	\item Appointment of honorary members.
	\item All other business.
	\end{enumerate}

	The General Assembly shall be considered valid regardless of the number of members present.

	Decisions of the General Assembly shall be taken by a majority vote of the members present. In the case of a deadlock, the President shall have the casting vote.

\english
	\subsubsection{Excutive Committee}\stepcounter{art}

	The Executive Committee is composed of at least 2 and at most 5 members. 
	All members of the Executive Committee are elected annually by the General Assembly.

	The Executive Committee is authorized to carry out all acts that further the purposes of the association. It represents the association, manages all issues and is responsible for all incomes and expenses and the financial status of the association. For special tasks, further persons may be involved by the Executive Committee.

	The Executive Committee holds a meeting at least once per month. 
	Decisions of the Executive Committee are considered valid if the majority of the Executive Committee present vote in favor of the decision. Votes can only be cast if the president is present. The president has the casting vote.

	New Executive Committee members will be selected by the Executive Committee and notified prior to the next General Assembly. Members can also recommend themselves to the Executive Committee. This must be communicated to the Vice President via email at least 3 months before the next General Assembly.

	Members of the Executive Committee must invest at least two hours per week on operational and organisational duties (not related to any particular project).

	The legally binding signature for the association is in hand of the president or vice president.

% =================================

\english
	\subsection{Finances}

\english
	\subsubsection{Resources}\stepcounter{art}
	The association funds its activities through

	\begin{enumerate}[(a)]
	\item sponsors
	\item membership fees
	\item donations
	\item events, workshops
	\item sale of goods and services
	\end{enumerate}
	\vspace{0.27cm}

\english
	\subsubsection{Liabilities}\stepcounter{art}
	Only the associations assets can be made liable for debts of the association. Personal liability of the Executive Committee or members is excluded.
		\vspace{1.5cm}

\english
	\subsection{Relationships with Third Parties}

	\subsubsection{EPFL}\stepcounter{art}
	The association respects the various rules concerning associations at EPFL.

	In particular, the association ensures that at least half of the Executive Committee is comprised of EPFL or UNIL members, the president being registered with EPFL or UNIL and that the associations activities target members of EPFL. 
			\vspace{2.5cm}

\english
	\subsection{Various}

\english
	\subsubsection{Revision of the Articles of \\ Association}\stepcounter{art}
	The present Articles of Association may be altered by the General Assembly if more than half of the members present and liable to vote agree. 		
	\vspace{0.5cm}

\english
	\subsubsection{Dissolving of the Association}\stepcounter{art}
	The association can not be dissolved if at least three members of the Committee vote for its continuing.
	All members must be notified by email in the case of the association being dissolved.
	Remaining assets should be transfered to another association with comparable purpose.

\english
	\vspace{\fill}

	\noindent
	Lausanne, 30.10.2016
	\vspace {1.5cm}

	\noindent
	\hrulefill \\

	\vspace {1.0cm}
 	\noindent 
 	\hrulefill \\

\clearpage
% ================================
% EMBED CODE OF CONDUCT

% \input{../code_of_conduct/code_of_conduct_content_english_french}

\end{paracol}

\end{document}